Stellar Bars in Isolated Gas-Rich Galaxies Do Not Slow Down

### 2022-04-06 paper/ms.tex
Move figures
Just move them all to the end
Add surface density
Adds the time evolution of the surface density

\begin{figure*}[b!]%
\centering
\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
\caption{Surface density projections in an N-body only simulation (upper
panels) and a simulation which includes the SMUGGLE model (bottom panels). We
can see that in the N-body run, the bar grows in length and strength. In the
SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
interstellar medium), the bar remains at approximately the same length and
strength over the course of the simulation. The N-body model is identical to
the GALAKOS model, discussed in the text.}\label{fig:overview}
\end{figure*}

@@

@@

\setcounter{page}{1}
\setcounter{figure}{0}
\setcounter{table}{0}
\captionsetup[figure]{labelformat=empty}
\renewcommand{\thefigure}{Extended Data \arabic{figure}}
\renewcommand{\thetable}{Extended Data \arabic{table}}

@@

The addition of the gas phase was done as follows. The version of
\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
disk which is radially exponential and in vertical gravito-hydrodynamic
balance. We modified the radial profile of this code in order to allow us to
generate a disk with a constant surface density within some radius, and then
exponentially declining beyond that radius with the scale-length of the
stellar disk. We used a surface density of $X$ and a maximum radius of $Y$. We
varied the surface density in TEST.

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS N-body disk (and bulge and dark matter halo) after
the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
purpose of allowing the GALAKOS to evolve first for a short period of time is
to allow for the bar to form unimpacted by the presence of the gas (which
would normally disrupt the formation of the bar). We made one additional
modification when stitching the gas disk together with the N-body disk - we
created a vacuum within the central $X$ pc. This vacuum guards against an
initial dramatic infall of gas within the bar region, which we found to
destroy the bar.

Our setup is initially out of equilibrium, but we found that after
$\sim300\,\textrm{Myr}$, the entire system has settled into a steady-state
configuration and initial transients appear not to affect the results after
this point. The constant surface density of the initial gas disk is important
for ensuring the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate. We show the evolution of the surface density
profile in Fig.~X compared to observational estimates (CITE).

>>>>

@@

\begin{figure*}[h]%
\centering
\includegraphics[width=0.8\textwidth]{fig/fig1.pdf}
\caption{Surface density projections in an N-body only simulation (upper
panels) and a simulation which includes the SMUGGLE model (bottom panels). We
can see that in the N-body run, the bar grows in length and strength. In the
SMUGGLE run (which includes a gaseous phase and a model for the multi-phase
interstellar medium), the bar remains at approximately the same length and
strength over the course of the simulation. The N-body model is identical to
the GALAKOS model, discussed in the text.}\label{fig:overview}
\end{figure*}

@@

\setcounter{page}{1}
\setcounter{figure}{0}
\setcounter{table}{0}
\captionsetup[figure]{labelformat=empty}
\renewcommand{\figurename}{Extended Data Figure}
\renewcommand{\thetable}{Extended Data \arabic{table}}

@@

The addition of the gas phase was done as follows. The version of
\texttt{MakeNewDisk} used for the original GALAKOS model can generate a gas
disk which is radially exponential and in vertical gravito-hydrodynamic
balance. We modified the radial profile of this code in order to allow us to
generate a disk with a constant surface density within some cut-off radius,
and then exponentially declining beyond that radius with the scale-length of
the stellar disk. We used an initial surface density of
$20\,M_{\odot}/\textrm{pc}^2$ and a cut-off radius of $9.3\,\text{kpc}$. We
varied the surface density in TEST.

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS N-body disk (and bulge and dark matter halo) after
the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
purpose of allowing the GALAKOS to evolve first for a short period of time is
to allow for the bar to form unimpacted by the presence of the gas (which
would normally disrupt the formation of the bar). We made one additional
modification when stitching the gas disk together with the N-body disk - we
created a vacuum within the central $4\,\textrm{kpc}$. This vacuum guards
against an initial dramatic infall of gas within the bar region, which we
found to destroy the bar.

Our setup is initially out of equilibrium, but we found that after
$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
configuration and initial transients appear not to affect the results after
this point. The constant surface density of the initial gas disk is important
for ensuring the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate. We show the evolution of the surface density
profile in Extended Data Fig.~\ref{fig:surf} compared to observational
estimates (CITE).

\begin{figure}[h]%
\centering
\includegraphics[width=0.45\textwidth]{fig/fig-surf.pdf}
\caption{The time evolution of the gas surface density (\textit{upper}) and the star formation rate (SFR) surface density (\textit{lower}).  } \label{fig:surf}
\end{figure}
